---
title: "Untitled"
author: "Vrinda"
date: "2024-10-12"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library('ggvis')
library('tidyverse')
library('ggplot2')
library('stats')
library('dplyr')
library('corrplot')
library('mosaic')

# read the file 
fast_food= read.csv("C:\\Users\\Vrinda\\Downloads\\FastFoodNutritionScraped20241011V2.csv", header=T, na.strings='')


```
**State the hypothesis**

1. Null Hypothesis(Ho) - Mean Trans Fat of the two companies is equal. i.e.$\mu{1}$ = $\mu{2}$

2. Alternative Hypothesis(H1) - Mean Trans fat of the two companies are unequal i.e.$\mu{1}$ $\neq$ $\mu{2}$
 




```{r}
#visualisation to see how trans fat is different based on compnay
violin_plot <- ggplot(fast_food, aes(x = Company, y = Trans.Fat..g., fill = Company)) +
geom_violin(trim = FALSE, draw_quantiles = c(0.25, 0.5, 0.75), alpha = 0.7) +
geom_boxplot(width = 0.1, fill = "white", color = "black", outlier.shape = NA) +
labs(title = "Violin Plot of Trans Fat by Restaurant",
x = "Company",
y = "Trans Fat",
fill = "Company") +
theme_minimal()
print(violin_plot)


#splitting the data based on company wise
company1_data <- fast_food$Trans.Fat..g.[fast_food$Company == 'A&W']
company2_data <- fast_food$Trans.Fat..g.[fast_food$Company == 'Dairy Queen']
company3_data <- fast_food$Trans.Fat..g.[fast_food$Company == 'Mcdonalds']
company4_data <- fast_food$Trans.Fat..g.[fast_food$Company == 'Subway']
company5_data <- fast_food$Trans.Fat..g.[fast_food$Company == 'Tim Hortons']
  


```
 AS the data is left skewed we can't use T-test as it is not normally distributed , so we are going with Bootstraping Distribution
```{r}
#bootstrap test as the data is not normally distributed
n = 1000
b = do(n)*(mean(resample(company1_data,replace = TRUE))- mean(resample(company4_data,replace = TRUE)))

quantile(b$result,c(0.025,0.975))
```
As we can see from the confidence interval we have the lower quantile in negative which implies that the mean Of Trans Fat is greater in Subway rather than A&W.
